---
title: "Diamond Price Analysis"
author: "Your Name"
output: html_document
---

# Introduction

This project analyzes the **ggplot2 diamonds dataset** to understand how diamond characteristics such as carat, cut, color, and clarity affect price. The analysis includes data exploration, visualization, correlation analysis, and linear regression modeling.

---

# Dataset Overview

The dataset used is the built-in `diamonds` dataset from the `ggplot2` package.

It contains:
- 53,940 observations
- 10 variables

Variables include:
- price
- carat
- cut
- color
- clarity
- depth
- table
- x, y, z (dimensions)

---

# Load Libraries

```{r}
library(ggplot2)
library(dplyr)
library(GGally)
library(corrplot)
```

---

# Load Dataset

```{r}
data("diamonds")
head(diamonds)
```

---

# Data Exploration

## Structure of Data

```{r}
str(diamonds)
```

## Summary Statistics

```{r}
summary(diamonds)
```

## Missing Values

```{r}
colSums(is.na(diamonds))
```

---

# Exploratory Data Analysis

## Price Distribution

```{r}
ggplot(diamonds, aes(price)) +
  geom_histogram(binwidth = 500, fill = "steelblue") +
  labs(title = "Distribution of Diamond Prices")
```

---

## Carat Distribution

```{r}
ggplot(diamonds, aes(carat)) +
  geom_histogram(bins = 30, fill = "darkgreen") +
  labs(title = "Distribution of Diamond Carat")
```

---

## Price by Cut

```{r}
ggplot(diamonds, aes(cut, price, fill = cut)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  labs(title = "Diamond Price by Cut")
```

---

## Price by Color

```{r}
ggplot(diamonds, aes(color, price, fill = color)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  labs(title = "Diamond Price by Color")
```

---

## Price by Clarity

```{r}
ggplot(diamonds, aes(clarity, price, fill = clarity)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  labs(title = "Diamond Price by Clarity")
```

---

## Carat vs Price

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Relationship Between Carat and Price")
```

---

# Correlation Analysis

```{r}
numeric_data <- diamonds %>%
  select(carat, depth, table, price, x, y, z)

cor_matrix <- cor(numeric_data)

corrplot(cor_matrix, method = "color", tl.cex = 0.8)
```

---

# Summary Statistics by Groups

## Average Price by Cut

```{r}
diamonds %>%
  group_by(cut) %>%
  summarise(AveragePrice = mean(price))
```

## Average Price by Color

```{r}
diamonds %>%
  group_by(color) %>%
  summarise(AveragePrice = mean(price))
```

## Average Price by Clarity

```{r}
diamonds %>%
  group_by(clarity) %>%
  summarise(AveragePrice = mean(price))
```

---

# Regression Model

We build a linear regression model to predict diamond price.

```{r}
model <- lm(price ~ carat + depth + table + x + y + z,
            data = diamonds)

summary(model)
```

---

# Prediction Example

```{r}
predict(model, newdata = data.frame(
  carat = 1,
  depth = 61,
  table = 57,
  x = 6.4,
  y = 6.4,
  z = 3.9
))
```

---

# Conclusion

This analysis shows that:

- Carat is the strongest predictor of price
- Price distribution is highly right-skewed
- Cut, color, and clarity all influence price
- A linear regression model can reasonably estimate diamond price
